---
title: "Data Tidying pt 2"
output: github_document
---

# Tidy Data Pt. 2 & Joins

This lecture is based on [Chapter 5 in R4DS](https://r4ds.hadley.nz/data-tidy) and [Chapter 19 in R4DS.](https://r4ds.hadley.nz/joins)

The last class on Tidying Data covered pivot_longer, which is what you're usually going to do to get data into tidy data format. Today we'll cover the opposite case, pivot_wider.

We'll spend most of the class talking about joins - different ways of combining multiple tibbles. This is really helpful conceptually for your requested topic next week: working with databases in R.

### pivot_wider()

Pivot wider creates columns and reduces rows - the opposite of pivot_longer(). The use case for pivot_wider() in the world of tidy data is to fix the problem of one observation being spread across multiple rows. 

```{r examine patient experience data}

```

In this case, we're looking at survey summary data. Instead of seeing the full data for one organization in each row, the data is spread across six rows. An analogy would be if every survey question for a respondent was in a separate row.

We can use the distinct() function (covered in DATA_3101_dplyr_pt1) to see the survey measures and related titles.

```{r distinct combinations}

```

Neither of these are great in terms of variable names. One carries no descriptive meaning and the other is a long text string including spaces. Think about what these would look like in a plot legend! We'll come back to this. Remember that Costanzo (2023) suggested that we structure the data before we clean it.

The structure of pivot_wider is the opposite of pivot_longer:

-   we provide the existing columns that define the values (values_from) 
-   where the column names are coming from (names_from)
-   which columns have values that identify the row (id_cols)

```{r pivot_wider}

```

Now we can work through some of the data cleaning and validation steps suggested by Costanzo (2023). In this case we can change the column names to something meaningful and check that the data types are correct.

This is also where we'd want to create a data dictionary for a study in a README.txt file, in terms of Research Data Management.

Let's take a look at the survey to better understand the variables: [CAHPS for MIPS Survey](https://www.cms.gov/data-research/research/consumer-assessment-healthcare-providers-systems/cahps-mips)

We could leave the column title as CAHPS_GRP_1, as the R4DS book does, or we could give it a somewhat meaningful name, like timeliness.

The entry in the data dictionary could look like:

CAHPS_GRP_1 = CAHPS for MIPS SSM: Getting Timely Care, Appointments, and Information, measured as a rate

or

timeliness = CAHPS_GRP_1: CAHPS for MIPS SSM: Getting Timely Care, Appointments, and Information, measured as a rate

Which would you use in a healthcare network where everyone new the codes and acronyms? What about at an international conference where people aren't familiar with this aspect of American health care?

## Joins

A dataset can include multiple tables. Sometimes in your analysis you might need to combine the tables.

There are two types of joins:

-   mutating joins: use to add new variables to one data frame from matching observations in another. Remember, mutate() adds new columns to a data frame.

-   filtering joins: filter the observations in one data frame based on whether or not they match an observation in another. Remember, filter() changes which rows are present.

Either way, you need to define the key, the variables that are used to connect the data frames. In the lecture by John Aspler on Tuesday, he talked about persistent identifiers like ORCID iDs and DOIs. These identifiers are excellent keys in all the joins that need to happen in the scholarly information infrastructure.

Today we'll go back to some of the tables in the nycflights13 package.

```{r load nycflights13}

```

#### Key definitions:

-   Primary key: variable or set of variables that uniquely identify each observation. It is called a compound key if more than one variable is needed.

-   Foreign key: variable or set of variables that correspond to the primary key in another table.

Ideally the primary key and the foreign key have the same names! Let's look at a [figure from R4DS](https://r4ds.hadley.nz/joins#fig-flights-relationships) showing the relationships between the nycflights13 data frames.

#### Check that keys are unique:

To verify that the primary keys are unique, you use count and filter to find out if there any duplicates. You can also check for missing values.

Note: You'd want to do this for airports (primary key: faa), weather (compound key: time_hour, origin), and airlines (primary key: carrier) as well.

```{r check for multiple entries and missing values}

```

#### dplyr joins: 

Mutating joins:

-   left_join(): keeps all rows in x

-   inner_join(): only keeps rows that occur in both x and y

-   right_join(): keeps all rows in y

-   full_join(): keeps all rows in both x and y

Filtering joins:

-   semi_join(): keeps all rows in x that have a match in y

-   anti_join(): keeps all rows in x that do not have a match in y

#### left_join()

left_join() is a mutating join. The main use case of left_join() is to add more metadata to the existing table. By default, left_join() uses all variables that are in both data frames as the join key.

**Task:**

**Add the full airline name to the flights dataframe**

```{r add full airline name}

```

Sometimes you will still need to specify the key, because you might have column names that mean different things in the different dataframes. For example, flights\$year is the year of the flight and planes\$year is the year the plane was built.

```{r}

```

#### Semi_join()

Semi_join() is a filtering join that keeps all rows in x that have a match in y.

**Task:**

**Filter the airports dataset to show just the origin airports in flights.**

```{r}

```

#### **Anti-joins()**

Anti-joins return all rows in x that don't have a match in y. You can use anti-joins to find implicit missing values.

Find which tailnums in flights are missing in planes

```{r tailnums not in planes}

```
